---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# paleval

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/paleval)](https://CRAN.R-project.org/package=paleval)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/ijlyttle/paleval.svg?branch=master)](https://travis-ci.org/ijlyttle/paleval)
[![Codecov test coverage](https://codecov.io/gh/ijlyttle/paleval/branch/master/graph/badge.svg)](https://codecov.io/gh/ijlyttle/paleval?branch=master)
<!-- badges: end -->

The goal of paleval is to help you evaluate the effectiveness of color palettes and color maps. It builds from the [colorspace](http://colorspace.r-forge.r-project.org) package, used to design color maps, and the [farver](https://github.com/thomasp85/farver) package, used to evaluate the perceptual difference between two colors.

This will have to go into a design-document eventually, but for now, I'll "puke" this here.

For the purpose of this package, following `colorRamp` and ggplot2 usage, we define some classes:

- **`pev_fcont`**: A function that describes a continuous palette. When called with a numeric vector with values between 0 and 1, it returns a vector of the corresponding (hex-code) values.

- **`pev_fdisc`**: A function that describes a discrete palette. When called with a single integer argument (the number of levels in the scale), it returns a vector of the (hex-code) value for the entire scale. There are two types of discrete-palette functions:

  - **`pev_bounded`**: Indicates that there is an upper-bound on the number of colors it can provide, e.g. [Tableau 10](https://www.tableau.com/about/blog/2016/7/colors-upgrade-tableau-10-56782). You would *not* think to interpolate between these colors.
  
  - **`pev_unbounded`**: Indicates that there is *no* upper-bound on the number of colors it can provide, e.g. [Tableau 10](https://www.tableau.com/about/blog/2016/7/colors-upgrade-tableau-10-56782). You might think to interpolate between these colors.

These follow the `palette` argument for [`ggplot2::continuous_scale()`](https://ggplot2.tidyverse.org/reference/continuous_scale.html) and [`ggplot2::discrete_scale()`](https://ggplot2.tidyverse.org/reference/discrete_scale.html), as well as `graphics::colorRamp()` and `graphics::colorRampPalette()`, respectively. 

I'm suspect I'm doing things in a way that the authors of `colorspace`, `ggplot2`, `farver`, and `graphics` might prefer not to consider while eating. As much as anything, this package is an attempt to reationalize and harmonize all of these concepts to myself. 

There are ways to create continuous-palette functions:

 * `pev_fcont()`: constructor
    * takes a vector (more than one) of hex-colors to be interpolated. 
    * takes a (one) name of a colorspace palette.
    * takes a set of HCL parameters, following the colorspace framework. 
    * takes a function.
    * takes a `pev_fcont` (no-op).
    
 * `pev_fcont_cvd()`: modifies the output of the palette function according to color-vision deficiency.
 * `pev_fcont_diverging()`: composes two continuous-palette functions to create (presumably) a diverging palette-function. 
   It's up to you to make sure the constituent palettes "meet in the middle". 
 * `pev_fcont_rescale()`: rescaling an existing continuous-palette function. It may make sense to "zoom-in", but "zooming-out" could get you into trouble. 
 * `pev_fcont_reverse()`: reverses the sense of the palette function.
 
There are ways to create discrete-palette functions:
 
 * `pev_fdisc()`: constructor 
   * takes a vector of hex-colors, returns a **bounded** function.
   * takes a `pev_fcont`, given a discretization `method` (e.g. `"panel"` or `"post"`), returns an **unbounded** function.
   * takes a `pev_fdisc` (no-op).
  
 * `pev_fdisc_cvd()`: modifies the output of the palette function according to color-vision deficiency.
 * `pev_fdisc_reverse()`: reverses the sense of the palette function.

 * `print()`: prints a representation of the palette function.
 
Other functions:

- `pev_nmax()`: get the maximum length supported by a discrete-palette function

* `pev_hex_distance()`: given two (sets of) hex-colors, `hex`, `hex_ref`:
    - return the perceptual `distance` between them 
    
* `pev_data_separation()`: assess color-separation, given a discrete-palette function,
    - return `data.frame` with `cvd`, `i`, `hex`, `hex_ref`, `distance`
* `pev_gg_separation()`

* `pev_data_derivative()`: assess perceptual-derivative, given a continuous-palette function,
    - return `data.frame` with `cvd`, `x`, `hex`, `d_distance_d_x`
* `pev_gg_derivative()`

* `pev_data_distance()`: assess perceptual-distance from a reference-color, given a continuous-palette function,
    - return `data.frame` with `cvd`, `x`, `hex`, `hex_ref`, `distance`
* `pev_gg_distance()`

- `pev_data_spectrum()`: assess HCL and RGB spectrum for continuous-palette functions, given a continuous-palette function,
    - return `data.frame` with `x`, `dimension`, `value`
- `pev_data_spectrum_hex()`
    - return `data.frame` with `x`, `dimension`, `value`, `hex_ref`
- `pev_gg_spectrum()`

* `pev_data_bloom()`: given `hex`, return `cvd`, `x`, `hex`, `hue`, `chroma`, `lumiance` 
* `pev_gg_bloom()`
 
## Installation

You can install the development version of paleval from [GitHub](https://github.com/ijlyttle/paleval) with:

``` r
devtools::install_github("ijlyttle/paleval")
```

<!--
You can install the released version of paleval from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("paleval")
```
-->

## Example

```{r}
library("paleval")

fcont <- pev_fcont("Dynamic") # continuous palette-function, from colorspace
fdisc <- pev_fdisc(fcont, method = "panel") # discrete palette-function

data_sep <- pev_data_separation(fdisc(7))
print(data_sep)

pev_gg_separation(data_sep)
```

```{r}
data_drv <- pev_data_derivative("Purple-Green")
data_drv

pev_gg_derivative(data_drv)
```


```{r}
data_dist <- pev_data_distance("Purple-Green")
data_dist

pev_gg_distance(data_dist)
```


```{r}
data_bloom <- pev_data_bloom("Viridis")
data_bloom

pev_gg_bloom(data_bloom)
```

```{r}
pev_data_bloom("Dynamic", n = 12) %>% pev_gg_bloom()
```

## Code of Conduct

Please note that the 'paleval' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
