---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# paleval

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/paleval)](https://CRAN.R-project.org/package=paleval)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/ijlyttle/paleval.svg?branch=master)](https://travis-ci.org/ijlyttle/paleval)
[![Codecov test coverage](https://codecov.io/gh/ijlyttle/paleval/branch/master/graph/badge.svg)](https://codecov.io/gh/ijlyttle/paleval?branch=master)
<!-- badges: end -->

The goal of paleval is to help you evaluate the effectiveness of color palettes and color maps. It builds from the [colorspace](http://colorspace.r-forge.r-project.org) package, used to design color maps, and the [farver](https://github.com/thomasp85/farver) package, used to evaluate the perceptual difference between two colors.

This will have to go into a design-document eventually, but for now, I'll "puke" this here.

For the purpose of this package, following ggplot2 usage, we define:

- **`pev_fcont`**: A (palette) function that, when called with a numeric vector with values between 0 and 1, returns the corresponding (hex-code) values.

- **`pev_fdisc`**: A (palette) function that, when called with a single integer argument (the number of levels in the scale), returns the (hex-code) values that they should take. Such a function will have a maximum for the argument.


These follow the `palette` argument for [`ggplot2::continuous_scale()`](https://ggplot2.tidyverse.org/reference/continuous_scale.html) and [`ggplot2::discrete_scale()`](https://ggplot2.tidyverse.org/reference/discrete_scale.html)

There are ways to create continuous-palette functions:

 - hcl, etc. parameters, following the colorspace framework. `pev_cont_hcl()`
 - composing two continuous-palette functions to create (presumably) a diverging palette-function. It's up to you to make sure the constituent palettes "meet in the middle". `pev_cont_compose()`
 - rescaling an existing continuous-palette function. It may make sense to "zoom-in", but "zooming-out" could get you into trouble. `pev_cont_rescale()`
 
There are ways to create discrete-palette functions:
 
 - discretizing a contiuous-palette function, specifying number of colors and discretization method (e.g. panels or posts). `pev_disc_cont()`
 - supplying a vector of hex-codes. `pev_disc_hex()`
 
As well, it would be useful to have a function that, given an `fpal_discrete`, returns the maximum-index.

## Installation

You can install the development version of paleval from [GitHub](https://github.com/ijlyttle/paleval) with:

``` r
devtools::install_github("ijlyttle/paleval")
```

<!--
You can install the released version of paleval from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("paleval")
```
-->

## Example

```{r}
library("paleval")

pal <- colorspace::qualitative_hcl(n = 5)
print(pal)

data_sep <- pev_data_separation(pal)
print(data_sep)

pev_gg_separation(data_sep)
```

## Code of Conduct

Please note that the 'paleval' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
